---
title: "P8105_ Homework 6"
Author: Katherine Dimitropoulou
Date: 11/25/2019
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 6,
  out.width = "90%")
library(tidyverse)
library(knitr)
library(readxl)
library(haven)
library(plotly)
library(viridis)
library(broom)
library(purrr)
library(modelr)
library(mgcv)
library(patchwork)
set.seed(1)
```
## Problem 1
Read in and clean birthweight data_ Clean variable names, drop missing data, recode variables to factor variables (i.e. babysex, race, etc.)
```{r, include = FALSE, warning = FALSE, message = FALSE}
birthweight_data = read_csv(file = "./birthweight.csv") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(babysex = as.factor(babysex),
         frace = as.factor(frace),
         malform = as.factor(malform),
         mrace = as.factor(mrace)) %>% 
  mutate(babysex = recode(babysex,
        '1' = "male", 
        '2' = "female")) %>% 
  mutate(mrace = recode(mrace,
        '1' = "white", 
        '2' = "black", 
        '3' = "asian", 
        '4' = "puerto rican",
        '0' = "other")) %>% 
  mutate(frace = recode(frace,
        '1' = "white", 
        '2' = "black", 
        '3' = "asian", 
        '4' = "puerto rican",
        '0' = "other",
        '9' = "unknown")) %>% 
  mutate(malform = recode(malform,
        '0' = "absent", 
        '1' = "present")) %>% 
  mutate_if(is.numeric, round, 2)
```

### Propose a regression model for birthweight.
The ouctome is birthweight (bwt).  
The tentative predictors are:  

Mother's Factors:  
- mrace: mother’s race  
- menarche: mother’s age at menarche  
- mheight: mother’s height  
- momage: mother’s age at delivery  
- ppbmi: mother’s pre-pregnancy BMI  
- ppwt: mother’s pre-pregnancy weight  

Gestation Factors:  
- gaweeks: gestational age in weeks  
- delwt: mother’s weight at delivery (pounds)  
- smoken: average number of cigarettes smoked per day during pregnancy  
- wtgain: mother’s weight gain during pregnancy  

Mother's Prior History Risk Factors:  
- parity: number of prior live births  
- pnumlbw: number of previous low birth weight babies  
- pnumsga: number of previous births with small for gestational age babies  

Other Family Factors:  
- fincome: family monthly income   
- frace: father’s race  

Baby Factors:  
- malform: presence of malformations that relate to weight  

Birthweight has been reported in the literature to relate to gestational age, history of previous low birth weight babies, mother's smoking status, income, and  malformations related to weight.  
##Model 1
```{r, warning = FALSE, message = FALSE}
birthweight_model_1 = lm(bwt ~ gaweeks + smoken + pnumlbw + fincome + malform, data = birthweight_data)
birthweight_model_1 %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3, caption = "Regression Summary (Model 1)")
```
## Residuals
```{r, warning = FALSE, message = FALSE}
birthweight_data %>% 
  modelr::add_residuals(birthweight_model_1) %>%
  modelr::add_predictions(birthweight_model_1) %>% 
  ggplot(aes(x = pred, y = resid)) + 
    geom_point(alpha = 0.5) + 
    geom_hline(yintercept = 0, color = "blue") + 
    labs(
      title = "Model Residuals vs. Fitted Values",
      x = "Fitted Values",
      y = "Residuals")
```

##Compare your model to two others:  
#Model 2: Using length at birth and gestational age as predictors (main effects only)  
```{r, warning = FALSE, message = FALSE} 
birthweight_model_2 = lm(bwt ~ gaweeks + blength, data = birthweight_data)
birthweight_model_2 %>% 
  broom::glance()  %>%
  knitr::kable(caption = "Regression Summary (Model 2)")
```

#Model 3: Using head circumference, length, sex, and all interactions (including the three-way interaction) between these  
```{r, warning = FALSE, message = FALSE}
birthweight_model_3 = lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = birthweight_data)
birthweight_model_3 %>% 
  broom::glance()  %>%
  knitr::kable(caption = "Regression Summary (Model 3)")
```

Make the comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.
```{r, warning = FALSE, message = FALSE}
cv_birthweight_models =
  crossv_mc(birthweight_data, 100) %>%
    mutate(train = map(train, as_tibble),
      test = map(test, as_tibble))
```

```{r, warning = FALSE, message = FALSE}
cv_birthweight_models = cv_birthweight_models %>%
  mutate(birthweight_model_1  = map(train, ~lm(bwt ~ gaweeks + bhead + blength + smoken + fincome + pnumlbw, data = .x)),
      birthweight_model_2  = map(train, ~lm(bwt ~ gaweeks + blength, data = .x)),
      birthweight_model_3  = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = as_tibble(.x)))) %>% 
    mutate(rmse_birthweight_model_1 = map2_dbl(birthweight_model_1, test, ~rmse(model = .x, data = .y)),
      rmse_birthweight_model_2   = map2_dbl(birthweight_model_2, test, ~rmse(model = .x, data = .y)),
      rmse_birthweight_model_3 = map2_dbl(birthweight_model_3, test, ~rmse(model = .x, data = .y)))
```

```{r, warning = FALSE, message = FALSE}
cv_birthweight_models %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
    geom_violin(aes(fill = model, color = model)) +
    labs(title = "Distribution of RMSE Values (Regression Models 1, 2, 3)",
      x = "Regression Model",
      y = "RMSE")
```

Model 1 has the lowest error (in prediction) followed by Model 3
## Problem 2 

Download Central Park Weather Data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

Focus on a simple linear regression with tmax as the response and tmin as the predictor:

```{r}
lm(tmax ~ tmin, data = weather_df) %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 3)
```

We are interested in the distribution of two quantities estimated from these data: r^2 and log(β^0∗β^1). Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities.

```{r}
weather_estimates =
weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x)),
    results = map(models, broom::tidy), 
    glance = map(models, broom::glance)) %>% 
  select(results, glance) %>% 
  unnest(results, glance) %>% 
  janitor::clean_names() %>% 
  select(term, estimate, r_squared) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate)  %>% 
  janitor::clean_names() %>% 
  mutate(
    log_betas = log(intercept * tmin)) 
```

Plot the distribution of your estimates.

```{r}
rsquared_plot = 
  weather_estimates %>% 
  ggplot(aes(x = r_squared)) +
  geom_density() +
  labs(
    title = "Estimation Distributions",
    x = "r-squared",
    y = "Density")
log_betas_plot = 
  weather_estimates %>% 
  ggplot(aes(x = log_betas)) +
  geom_density() +
   labs(
    x = "log(β0*β1)",
    y = "Density")
(rsquared_plot + log_betas_plot)
```
*These plots demonstrate that the estimates are approximately normally distributed.*

Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r^2 and log(β^0∗β^1).

```{r}
pull(weather_estimates, r_squared) %>% 
  quantile(c(0.025, 0.975))
pull(weather_estimates, log_betas) %>%
  quantile(c(0.025, 0.975))
```

*The 95% CI for r-squared is (0.8935, 0.9266). The 95% CI for log(β^0∗β^1) is (1.9654, 2.059).* 

